SHARP - II: Revealing a bias in observational measurements of dark matter substructure with gravitational lens flux ratios

### 2016-02-09 papers/B1555_EdgeOn/Resubmission_v2/referee_response_2.txt

>The revised version of the paper still contains the unsubstantiated speculation pointed out in my previous report, with no additional supporting evidence.  I am puzzled by the authors' statement in their response that an analysis of the effects of disks on lens statistics would be beyond the scope of the paper, since this is literally the topic of the paper.  The main problem now is that the paper's title is misleading, since nothing in this work provides any reason to believe that there is a bias in observational measurements of dark matter substructure with gravitational lens flux ratios.

RESPONSE: The referee is correct to point out that our title may be misleading in implying that we intend to discuss the effects of disks on lens statistics. Thus, we add more description on how can we be informed by a new model of a single system in the discussion. That being said, discs with high inclination in lens system can significantly affect the lensing cross-section, which has been shown in Maller et al. 1997, Möller et al. 2003, and Keeton & Kochanek 1998.  Although the fraction of high-inclination spiral lenses may not be high in the general population, the current sample of lensed quasars that have flux-ratio anomalies is fairly small size.  Therefore, if even a few systems within the sample have edge-on discs, these are sufficient to bias the substructure analysis. As the current work of flux anomalous lenses are analyzed with a standard SIE mass profile, it is fair to say that the disc effect is a bias that been ignored for a long time. Moreover, from the high-resolution IR imaging among these flux anomalous lenses, we know that B1555 is not the only system having an edge-on disc across lensed images. Since this is the first time an edge-on disc been confirmed that can be accounted for, at least, part of flux anomalies, we think it is reasonable to keep our title with the wording ‘bias’ rather scope down the implication from our results. [But to be more conservative about our finding on disc-effect, we decide to rephrase 'bias' into 'potential bias' in the title. ]

As the reason describing above, we seem to have a fundamental disagreement with the referee about the topic of the paper.  The referee feels that it should be an analysis of the statistical effect of disks in lenses.  Most/all previous substructure analyses have interpreted _all_ radio-wavelength flux ratio anomalies as being due to substructure.  We feel that the point of this Letter is to show that the previous assumptions may be wrong, since we have a clear case in which a disc _may_ produce the observed flux-ratio anomaly.  Having made that point in this short letter, we intend to write (and are actively working on) a longer paper that does a full statistical analysis.

@@

>As noted in my previous report, the observational result is quite interesting and merits publication. Since the authors were unable to support their speculations about biases in lens modeling, and have revised the text to focus more on this one specific lens system, I recommend that they amend the title and abstract to more accurately reflect the actual contents of the paper. Something along the lines of "An edge-on disk in the anomalous lens B1555+375" would be more accurate and less misleading than the current title, which is unacceptable for publication given what is actually in the paper.  Similarly, the last 2 sentences of the abstract is misleading.  Although the authors' observations lack sufficient signal/noise to constrain the amplitude of the disk mass in this system, in other lens systems with arcs, the fraction of mass in disky components is constrained to be <~ 1% of the mass in the Einstein radius.  Unless the authors present evidence or arguments for why these previous findings are incorrect, the last sentence of the abstract should be removed.

RESPONSE: We understand that if a constraint from photometry can be applied on B1555+375, it will help us to claim that the best-fit model we found is a ’true’ model. However, the goal of this paper is to demonstrate and bring the attention to the community about the disc-effect. Indeed, as the referee mentioned, the disc mass fraction in the Einstein radius can be down to  <~1%. This is true for massive elliptical lenses but does not hold for spiral lenses especially for lenses having high-inclined discs. SWELLS lenses, a group of spiral lenses selected from SDSS spectra, provide clear evidence of high fraction of stellar mass within the Einstein radius. Figure 5 in Brewer et al. 2012 shows that stellar mass dominates the lens mass. This is quite common in the small-seperation lens systems since that the very inner region of galaxies are mostly stellar mass. And these systems are more likely to be spirals. By fitting to photometry, we conclude that B1555+375 is a late-type spiral with no appreance of bulge. We think it is clear that saying the disc of B1555+375 accounts for less than 1% of lens mass is unphysical base on the stuides of spiral lenses.

Moreover, by combining the information given in Treu et al. 2011 and Brewer et al. 2012, the disc mass fraction of SWELLS lenses ranges from 5% to 30% under a conservative 'minimal' Chabrier ISM assumption. Especially when we look at J0820+4847, being a good anologous system to B1555+375, we notice that it has an arc right across its edge-on disc and has 15% of disc mass fraction within the Einstein radius. Again, the studies on spiral galaxies support our model having ~15% of disc mass fraction. The cases hold for disc mass fraction <~1% is most likely massive elliptical lenses, but not edge-on spiral lenses.

### 2016-02-09 papers/B1555_EdgeOn/B1555_letter.tex
little change to B1555 text

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:3panel}c. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the \textbf{ExpDisc component makes up about 15 per cent of the total projected mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total projected mass. The sample of 16 edge-on disks from SDSS found in the SWELLS survey shows that the disc can contribute more than 20 per cent mass in the Einstein radius \citep{swells1,swells3}.} 
%In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. 
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.

>>>>

The best-fit model parameters and 68\% CL errors are listed in Table~\ref{tab:model} and the lens model is illustrated in Fig.~\ref{fig:3panel}c. Table \ref{tab:results} also shows the comparison between the observed radio data and the model-predicted image positions and flux ratios. We find that by adding an elongated exponential disc to a smooth SIE model there is significant improvement in the predicted flux ratios when compared to the results for the SIE-only model. 
% with a reduced $\chi^2 = 2.7$
The positions are all in agreement, and the flux ratios are consistent at the $2\sigma$ level, with excellent agreement now obtained for the two merging images A and B. The remaining differences in fluxes may due to the real lensed source being slightly extended rather than the point source that we used in our model. The model requires that the \textbf{ExpDisc component makes up about 15 per cent of the total projected mass within the Einstein radius (1.7~kpc for an assumed lens redshift of $z_l = 1$).  In this central region of an edge-on spiral galaxy, the baryonic mass can be a significant component of the total projected mass. Studies on edge-on spiral lenses, for example the SWELLS survey, show that the disc can contribute more than 20 per cent mass in the Einstein radius \citep{swells1,swells3}.} 
%In the NIR imaging there is no sign of a bulge; a bulge plus disk fit to the lens galaxy surface brightness distribution gives a bulge component that is consistent with zero emission.  We therefore interpret the ExpDisc component as a good description of the baryonic mass of the lens galaxy. 
%We note that there is a significant 0.036~arcsec offset between the SIE and ExpDisc components, which is likely due to an additional ellipticity in the model (e.g. from an external shear) that is being partly compensated for by the disc component. Again, more complex mass models require the stronger constraints provided by deep imaging of the full extent of the gravitational arc in future high resolution radio observations.
